---
title: "Warnecke_Week_4_Random_Forests_Assignment"
author: "Aaron Warnecke"
date: "2/16/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries and Setup
```{r, include = FALSE}
library(tidyverse)
library(tidymodels)
library(ranger) #for random forests
library(caret)
library(gridExtra)
library(skimr)
library(vip) #variable importance
```

Read in data and add names to columns
```{r}
drug = read_csv("drug_data-1.csv")

names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
                "Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
                "SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
                "Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
                "LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

#str(drug)

```

Change usage classification to binary
```{r}
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"
```

Further cleaning and recoding
```{r}
drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
                mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
                                                    "45_54", "55_64", "65_"))) %>%
                mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
                mutate(Education = factor(Education, labels =
                  c("Under16", "At16", "At17", "At18", "SomeCollege",
                    "ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
                mutate(Country = factor(Country,
                labels = c("USA", "NewZealand", "Other", "Australia",
                            "Ireland","Canada","UK"))) %>%
                mutate(Ethnicity = factor(Ethnicity,
                labels = c("Black", "Asian", "White", "White/Black", "Other",
                           "White/Asian", "Black/Asian"))) %>%
                mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
                select(-ID)

#str(drug_clean)
```

Focus on Nicotine only
```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

## Task 1
```{r}
skim(drug_clean)
```
No missingness here! Got lucky.

## Task 2
```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) # 70% in training
train = training(drug_split)
test = testing(drug_split)
```

## Task 3
Visualization  
```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
grid.arrange(p1,p2,p3,p4)
```

*Age:* Nicotine use decreases with increasing age
*Gender:* More males than females use Nicotine
*Education:* This one is interesting. Under 16 had the most Nicotine use, while 16 years old cuts down quite a bit. Then Nicotine increases again from 17 to 18 and then to Some College. Goes down with more education after that like you would expect.
*Country:* Ireland and the USA are the heaviest users out of all countries in the study. The UK is the least.

```{r}
p1 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```

*Ethnicity:* White/Black and White/Asian have the highest Nicotine usage, where pure Black and Asian have the lowest.
*Nscore:* Higher neuroticism lends to higher Nicotine use.
*Escore:* Extraversion has little effect on Nicotine use.
*Oscore:* Openness to experience leads to higher Nicotine use.

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4)
```

*Ascore:* Agreeableness is inverted from what I would expect: the more agreeable you are, the less likely you are to use Nicotine.
*Cscore:* Lower conscientiousness leads to more Nicotine use, although not a strong correlation.
*Impulsive:* Higher impulsiveness correlates to higher Nicotine use.
*SS:* Higher sensation seeing leads to higher Nicotine use.

## Task 4
Set up our folds for cross-validation  
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```

Random forest with an R-defined tuning grid
```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% # add tuning of mtry and min_n parameters
  set_engine("ranger", importance = "permutation") %>% # added importance metric
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)),
  min_n(range = c(5, 20)),
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid
)
```

Visualization
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

This plot doesn't seem to tell us much about min_n and accuracy. Accuracy seems to peak at mtry or 3, then decreases with increasing mtry.

An alternate view of the parameters:
```{r}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```

Likewise, min_n still seems to be scattered. min_n of 5 and mtry of 3 seem to give us our best result from this plot.

## Task 5
```{r}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
Looks like I misinterpreted the pinks from the 2nd graph - min_n of 20 was our best.

```{r}
# fit the finalized workflow to our training data
final_rf_fit = fit(final_rf, train)
```

Check out variable importance
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

Sensation seeing is the most important variable by far. Being from the UK and openness to experience are the next moderately important.

## Task 6
Predictions  
```{r}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)
```

Confusion matrix
```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```

Predictions on test
```{r}
testpredrf = predict(final_rf_fit, test)
#head(testpredrf)
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

The model gives us 84.4% accuracy on the training set and 71.2% accuracy on the test set. The test accuracy is significantly lower than the training, but not low enough to cause too much concern.

## Task 7
In this example, we looked at socioeconomic factors and how they affect Nicotine use, but there are many other drugs included in this data set. One use for this model would be to predict how likely a high school or college student is to fall prey to drug abuse based on their personalities and demographics, and take preventative measures. This model's accuracy was not super high, but good enough perhaps to target certain students (such as a school, county, etc) for drug education.












